### 2015-10-01 realImplementation.tex

\section{Real hardware implementation}
The implementation of CSMA/ECA$_{\text{Hys+FS}}$ should be carried out at a firmware level due to the tight timing constrains related to the backoff procedure. There are several alternatives to modify the backoff operation in WNICs, either using open firmware like OpenFWWF~\cite{OpenFWWF} or through Field Programmable Gate Arrays (FPGA).

Although the FPGA option would in theory provide the strict timing requirements, the associated costs and difficulty are greater than the OpenFWWF alternative. OpenFWWF provides an open CSMA/CA firmware for a specific model of Broadcom WNICs' chipset, so the resulting firmware can be uploaded and tested in real commercial hardware.

By making the modifications represented in Algorithm~\ref{alg:CSMA_ECA}, a proof-of-concept prototype of CSMA/ECA was built using comercial hardware~\cite{ECA-DEMO-INFOCOM14, sanabria2013prototyping, BECA-test}. The prototype was able to create a collision-free schedule among a reduced number of contenders using big deterministic backoffs (such as $128$ or $256$ slots). Nevertheless, clock drift, imprecise timing~\cite{bianchi2007experimental} and other prototyping non-idialities prevented the construction of a collision-free schedule when using a small deterministic backoff or dynamic rate adaptation algorithms, like Minstrel~\cite{minstrel}.

Although the timing inaccuracy issues were identified and leveraged to prototype a similar protocol in real hardware~\cite{CF-MAC}, further development on the prototype is being carried out in order to incorporate Hysteresis and Fair Share.