### 2015-10-01 pendingChanges.tex

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{mathtools}
\begin{document}

\title{Fixes to the original paper}
\author{Luis Sanabria-Russo}
\date{Today}
\maketitle

\section{Abstract}
\begin{enumerate}
	\item Include remarks of the real hardware implementation.
\end{enumerate}

\section{Introduction}
\begin{enumerate}
	\item While describing the lack of studies of CSMA/ECA in non-saturated scenarios, you should also talk about the implementation on real hardware and the introduction of channel errors.
\end{enumerate}

\section{Related Work}
\begin{enumerate}
	\item We should introduce the performance evaluation of L-ZC and L-MAC with Basic CSMA/ECA/
\end{enumerate}

\section{CSMA/ECA}
\begin{enumerate}
	\item Figures do not appear as they are mentioned. For instance, Figure 2 appears before Figure 2.
	\item A reviewer suggests dividing the figure to highlight which curves are associated with which axis. 
	\item Arrange the legend of the figure to mimic the order of appearance of the curves. From top to bottom.
	\subsection{Supporting many more contenders}
	\begin{enumerate}
		\item Although we only treat CSMA/CA with maximum aggregation, here we should incorporate CSMA/CA$_{\text{FS}}$.
	\end{enumerate}
\end{enumerate}

\section{Simulation Scenario}
\begin{enumerate}
	\item We should mention the channel errors scenario, too.
	\item The reference to the simulator should be changed to the one we are actually using.
	\item We should take a moment to explain why we are not using RTS/CTS and the hidden nodes assumption.
	\item Table I should be reviewed.
\end{enumerate}

\section{Results}
\begin{enumerate}
	\item We should include the channel errors scenario alongside the non-saturated scenario.
	\item Coexistence figures should be focused on the per-station statistics.
\end{enumerate}

\section{Real Hardware Implementation}
\begin{enumerate}
	\item Should it be a subsection of the results?
\end{enumerate}

\end{document}